\section{Birth Certificates}

American law generally holds the naming of children to be the right and
responsibility of parents, without shutting the door on regulating edge cases,
like "Ghoul Nipple", "Legend Belch",
"Brfxxccxxmnpcccclllmmnprxvclmnckssqlbb11116", and "" \parencite{larson11}. In
many American states, however, this right is abridged with reference to
diacritical marks above letters, hardly an edge case in many languages around
the world. \textcite[5]{larson11} investigates this in his study of American naming
law, finding states with such rules to include California, Massachusetts, New
Hampshire, and Kansas:

\subsection{California}

Guidelines provided by the California Office of Vital Records (OVR) instruct
county clerks that baby names may contain only "the 26 alphabetical characters
of the English language with appropriate punctuation if necessary" and that "no
pictographs, ideograms, diacritical marks (including 'é,' 'ñ,' and 'ç') are
allowed" \parencite{larson11}.

The OVR's handbook cites Proposition 63, a 1986 ballot referendum which declared
English the state's official language, with the support of 74\% of voters. The
initiative created Article III, Section 6 of the California Constitution, which
not only explicated the status of English, but also entrusted to the state
government broad powers of enforcement: 

\begin{aquote}{California Constitution, Article III, Sec. 6(c)}

The Legislature shall enforce this section by appropriate legislation. The
Legislature and officials of the State of California shall take all steps
necessary to insure that the role of English as the common language of the State
of California is preserved and enhanced. The Legislature shall make no law which
diminishes or ignores the role of English as the common language of the State of
California.

\end{aquote}

The California Department of Public Health may see the prohibition of
"non-English" characters as a natural result of Proposition 63; other government
entities in California interpret the law differently. Two California state
parks, Año Nuevo State Park and Montaña de Oro State Park, manage to contain the
Spanish ñ in their official names, which is reflected on the parks' official
webpages. \parencite{año-nuevo} \parencite{montaña-de-oro}

Likewise, the City of San José, California includes the accented é in its
official name, and its Style Guide includes instructions on how to produce it
digitally: "To create an accented é, hold down the alt key and type '0233'‚ on
the numeric key pad." California's Department of Public Health likely disobeys
the city's guidelines in birth certificates, though this needs to be verified.

\subsubsection{2014-AB-2528}

A 2014 bill in the California State Assembly sponsored by AM Nancy Skinner
(AB-2528) sought to rectify the state's processing of birth certificates and
driver's licenses by allowing diacritical marks in names. The bill "required the
State Registrar to ensure that diacritical marks on English letters are properly
recorded on birth certificates, death certificates, certificates of fetal death,
and marriage licenses, including, but not limited to, accents, tildes, graves,
umlauts, and cedillas". \parencite{ab-2528}

AB-2528 stalled in the Appropriations Committee once state agencies assigned
multi-million dollar price tags relating to IT upgrades, noting that the DMV's
software could not "even accept lower-case letters". For this same reason the
bill was opposed by the County Recorder's Association of California.

\subsubsection{2017-AB-82}

In 2017, California AM Jose Medina revived the issue with AB-82, which
ultimately passed both houses of the legislature before being vetoed by Governor
Jerry Brown. Unlike the 2014 bill, this edition did not affect the issuance of
driver's licenses, only birth certificates. Passing through many more stages of
the legislative process, the committee hearings gathered more detailed estimates
for the cost of IT upgrades than they had in 2014:

\begin{itemize}

\item \$230,000 for IT upgrades at Department of Public Health
\item \$2 million per year for Department of Public Health to correct existing
records
\item Loss of revenue of \$450,000 per year to Department of Public Health since
they would not be able to electronically transmit names to SSA (at \$3 per name)
containing diacritics
\item Up to \$12 million for local governments to upgrade their systems
\item \$1--3 million in upgrades to Department of Health Care Services
\item Unknown administrative costs to Department of Social Services

\end{itemize}

The sticking point for Governor Brown was compatibility with federal databases,
which do not accept diacritics. In his veto message, he argued that the risks to
vital records outweighed the benefits of cultural openness:

"Mandating the use of diacritical marks on certain state and local vital records
without a corresponding requirement for all state and federal government records
is a difficult and expensive proposition. This bill would create
inconsistencies in vital records and require significant state funds to replace
or modify existing registration systems."

The committee findings make clear that the state would incur nontrivial costs to
update the name registration systems. Little discussion is included of the
possible creative solutions to the problem. Even assuming that government
systems cannot be made to support the full UTF-8 standard, there are ways of
representing information using ASCII. For example, we will see later that the
international specification for machine-readable passports has a variety of
control sequences for representing subtle distinctions in the Latin, Cyrillic,
and Arabic alphabets using only the 26 plain characters of the English alphabet.
The original form can be recovered nearly losslessly using the transliteration
table.

\subsection{Massachusetts}

In Massachusetts, the "characters have to be on the standard american keyboard.
So dashes and apostrophes are fine, but not accent marks and the such"
\parencite{larson11}.

\subsection{New Hampshire}

"All special characters other than an apostrophe or dash" are prohibited
\parencite{larson11}. Technical limitations of the state's database systems
prevent the inclusion of any diacritical marks.

\subsection{Kansas}

Restrictions are similar to those in Massachusetts \parencite{larson11}.
