\section{Hawaii}

Hawaii is the only state in the United States having two official languages, English and Hawaiian.

\subsection{The Case of Janice "Lokelani" Keihanaikukauakahihuliheekahaunaele}

A Hawaiian woman named Janice Keihanaikukauakahihuliheekahaunaele has faced some
administrative hassle due to her lengthy surname. She made the international
press in 2013 for her crusade against Hawaii's Bureau of Motor Vehicles,
demanding an ID card which would include her full name. The recent renewal card
she had received omitted her first name entirely, including her surname, but
with the final 'e' chopped off. This caused for her an awkward situation at a
traffic stop, where the policeman questioned her lack of a given name. After
some complaining to the authorities, the state was able to extend the the
character limit from 35 to 40 characters and issue Ms.
Keihanaikukauakahihuliheekahaunaele a revised driver's licence.

To the bureaucrats handling her case, she was presumably an annoyance to whom
their first suggestion would be to change the name, perhaps back to her maiden
surname. Unfortunately Ms. Keihanaikukauakahihuliheekahaunaele did not
appreciate this suggestion, whose surname serves as a link to her late husband,
whose full name was Keihanaikukauakahihuliheekahaunaele, a traditional Hawaiian
name carrying genealogical information to those few who still understand
traditional Hawaiian language (supposedly cherished by the state). As she wrote
in an email to a tool company:

\begin{aquote}{\parencite{lee-valley}}
A majority of people we attempt to do business with do not take the time to
	spell out the whole name and we do not do business with them again. We
	appreciate the respect you have shown us and will do business with you again.

You see, to some people in the world, your name is everything. If I say my name
	to an elder Hawaiian (kupuna), they know everything about my husband's family
	going back many generations... just from the name. When the name is sliced up,
	changed or altered it distorts the intention and meaning that the name
	represents. Unfortunately, many people have been shamed into hiding their real
	names because they don't fit in with the dominant culture's lack of respect
	for the name.

P.S. If bills or traffic citations are not correctly addressed, my husband
	refuses to pay and is under no obligation legally to pay.
\end{aquote}

An arbitrary character limit, such as 35, is result of the importance of
databases in managing government and corporate information; in many database
systems, the number of characters allotted for a text field must be specified
beforehand in the database schema. A strict character limit for names is almost
nonsensical in the absence of a rigid data-storage format. Presumably, the
state's official recognition of Hawaiian was partly responsible for ensuring a
favorable response to her request; we can imagine the issue quickly escalating
from a media curiosity into a genuine debate about Hawaii's commitmment to
language and racial justice. We should contrast the successful outcome of Ms.
Keihanaikukauakahihuliheekahaunaele's case with the uncoöperative bureaucracy in
China regarding the names of Ma Cheng and Zhao C.
