\section{Iceland}

Iceland's naming system is distinctive for retaining its patronymic (non-fixed)
surnames and the government's maintenance of a Personal Names Register for
child-naming. Both policies reflect an insular nationalism and Icelander's
pride for their conservative North Germanic language. Like Latvia, the
Icelandic state actively promotes its language occurs in a postcolonial
context, asserting its unique heritage out from the linguistic shadow of Danish
rule. Controlling the spread of modern-type fixed surnames, which threatened to
displace the indigenous patronymic system, was seen as a way of maintaining
Icelandic identity \parencite{willson02}. The more recent institution of the
Personal Names Register also represents an effort to maintain the cultural and
grammatical structure of Icelandic. However, the appropriateness of Icelandic
regulations should be considered from the human rights perspective as well,
especially in the wake of the UN decision in Latvia discussed above.

\subsection{Icelandic nationalism and purism}

Nationalist and Romantic tendencies gained strength in Iceland in the late
nineteenth century in the context of Iceland's struggle for independence from
Denmark. Icelanders claimed their unique linguistic heritage as one
justification for political independence. The ideology of linguistic purism
went virtually unchallenged within the general populace.
\textcite{kristinsson12} notes that there was harmony between "practices,
beliefs, and management decisions" in language policy during Iceland's period
of peak nationalism between 1860 and 1960.

Icelandic linguistic policy shifted focus in the 1960s, when the state created
the Icelandic Language Council. The postwar presence of US forces at a
permanent military installation marked the symbolic beginning of Anglophone
cultural imperialism. \textcite{kristinsson12} interprets the creation of the
Council as a reaction to the more fragile position of Icelandic and waning
linguistic nationalism, as linguistic policymakers sought to stem the tide of
global cultural influence on Icelandic.

\subsection{Surnames debate}

Although Icelandic faced an unprecedented global threat in the aftermath of
World War II, we should note that this phenomenon was foreshadowed around the
turn of the century in the debate over whether Iceland should adopt modern
fixed surnames, as most of Europe had already done. Even during this time,
linguistic purism was not maintained without any legal intervention. One of
Icelandic's salient characteristics is the patronymic system of surnames. The
majority of Icelanders take as a surname their father's or mother's given name,
in the genitive case and with the appropriate "-son" or "-dóttir" suffix
attached, depending on gender.

\parencite{willson02} notes that the preservation of this unique system was not
inevitable, and policymakers and linguists at the turn of the century debated
whether Iceland should adopt fixed surnames. Despite the linguistic purism of
much of Iceland's population, this debate was responding to a real surge of
Continental-style surnames in the nineteenth century, as persons from the
upwardly mobile class pursued studies in Denmark or elsewhere in Europe and
brought the practice home to Iceland. Ultimately the spread of surnames was
only prevented via legal action.

The first attempt to limit surname adoption was an 1881 bill introduced in the
Alþingi (Icelandic legislature) instituting a tax on the adoption of such
names. The proposal would have required royal permission to adopt a surname,
and, if approved, would require the payment of five hundred krónur; hereafter
the nameholder would also be subject to an annual title tax
("nafnbótarskattur") of ten krónur per syllable in the name
\parencite[137]{willson02}. This proposal was not approved, however, and the
debate raged on about what should be done.

To briefly summarise the debate presented in-depth by \textcite{willson02}, we
note that the two camps roughly represented the modernist and conservative
perspectives. Writer Guðmundur Kamban presents an internationalist defence of
surnames (he himself used a fixed surname), claiming that female Icelanders
abroad will inevitably be referred to by their husband's patronymic, creating a
gender mismatch. Moreover he praises the extreme morphological simplification
of English as a sign of linguistic development. Countering Guðmundur's
forward-thinking approach, the conservative line was carried by linguist
Jóhannes Jóhannsson, who touted the linguistic heritage of Icelandic and
suggested that, should surnames prove inevitable, they should be made more
harmonious with the language by incorporating authentic Icelandic elements.

A law on the topic finally passed the Alþingi in 1913, levying a moderate
charge for surname adoption without any annual tax. A government report in
1915, \textit{Íslenzk mannanöfn} (Icelandic Personal Names), carried on the
work of Jóhannes in trying to develop an Icelandic adaptation of modern
surnames and went to great lengths to justify their recommendations with sound
philological argumentation. Still, the report emphasised the need for
modernisation, as the writers chose not to include any surnames including the
letter Þ, due to its likelihood for corruption in foreign contexts.

Apparently this report stretched the linguistic purism of Iceland's population
to the breaking point. The report's suggestions were ridiculed by writers as an
indication that surnames were inherently incompatible with Icelandic. The
counterreaction was consummate in 1925, when the adoption of new surnames was
firmly forbidden by the Alþingi. Although ultimately the patronymic system
remained intact, the Icelandic surname debate casts doubt on the inevitability
of maintaining Icelandic's purity. Legal intervention was necessary to control
the public's naming practices and prevent a structural change of the national
language.

\subsection{Personal Names Committee}

A better-known aspect of Iceland's naming policy is the Personal Names
Committee (Mannanafnanefnd). The committee of three people maintains a registry
of permissible forenames that are "capable of having Icelandic genitive
endings", "written in accordance with the ordinary rules of Icelandic
orthography", and gender-conforming: "girls shall be given women's names and
boys shall be given men's names" \parencite{isl-names}. If a parent wishes to
use a name absent from the list, they must petition the Committee for approval.
Along with their ruling (which is explicity non-appealable under the Personal
Names Act), the Committee publishes an explanation of their decision. As a
typical example of their reasoning, "Huxland" was rejected in 2014 because
there was no precedent for geographical suffixes in given names (only in middle
names!) and because the suffix "-land" implies a neuter noun
\parencite{kyzer15}.

The most famous case from the Naming Committee is that of Blær Bjarkardóttir
Rúnarsdottir, whose first name was rejected by the Committee, since they
classified it as a male name. However, the family refused to back down and
choose a new name. Thus the state applied the generic forename Stúlka ("girl")
to all of Blær's official documents, including her passport. Iceland takes
enforcement of naming policies seriously. Only after arguing that the name Blær
was used for a female character in a popular Icelandic novel did Reykjavík
District Court allow the name to be used.

\subsection{Human rights implications}

Iceland's practice of labelling children as "Stúlka" (girl) or "Drengur" (boy)
if their parents do not comply with the Personal Names Act raises concerns
under human rights, as this could be considered demeaning and disproportionate
to the goal of supporting an official language. A generic placeholder name is
arguably no name at all, and this would constitute a violation of Article 7 of
the Convention on the Rights of the Child, which states that all to children
have the right to a name and to which Iceland is a signatory \parencite{crc}.
This being the most bare-bones conception possible of naming rights, Iceland's
violation should be addressed by the international community.

In the light of the United Nations Human Rights Committee's decision in
\parencite{raihman10}, it is difficult to see how Iceland's policy, if
challenged in this forum, would not be struck down likewise. Like the case
in Latvia, Iceland's Personal Names Act circumscribes individuals' choice of
names in order to support the official language. Recalling that Article 17 of
the International Covenant on Civil and Political Rights encompasses the "right
to choose and change one's own name", Iceland's policy contravenes the state's
obligations under this provision. Despite Iceland's legitimate interest in
protecting their national language, it is imperative that linguistic management
does not interfere with individual freedoms.
