\documentclass{article}

\usepackage[margin=1in]{geometry}
\usepackage{comment}
\usepackage{fontspec}
\usepackage{defreitasgabe}
\usepackage[backend=biber]{biblatex}
\usepackage[english]{babel}
\usepackage[autostyle,english=american]{csquotes}

\MakeOuterQuote{"}
\addbibresource{Bibliography.bib}
\setmainfont{Georgia}

\begin{document}

\section*{Introduction}

Although the United States has no official language specified in the US
Constitution, one might easily assume that English holds this title, considering
its ubiquitous role in American public life. A majority of US states have indeed
enshrined the official status of English through constitutional or statutory
means. State language laws are the primary object of study in this project.

Sociolinguistics affirms that language does not exist in some theoretical
vacuum; it carries social evaluations and entails real outcomes for real
speakers of a particular language variety. Through a perspective of political
linguistics, we will deconstruct the linguistic policy of the United States to
understand how a multicultural nation ultimately abetted the hegemonic status of
English in the US. We will look at how social evaluations and their legal
manifestations played out over the history of the United States, and the effects
(or lack thereof) of laws on US language communities.

I want to investigate the current state-level status of English as an official
language, group the laws by their forms and effects, and locate reliable data
that will help gauge the efficacy of language laws and compare them to other
measures taken in multilingual countries.

\section*{Consequence}

The percentage of Americans who speak a language other than English as a first
language continues to rise. Some states, such as California, with Official
English laws on the books, have large numbers of minority-language speakers
whose perspectives must be considered vis-à-vis the majority language. In these
cases, language issues directly reflect the struggle for political power between
majority and minority in a democratic society.

Furthermore, the election of Donald Trump as President has alerted public
consciousness to renewed fears over immigration and the presence of
foreign-language speakers in the US. \textcite{Liu14} finds that the national
salience of immigration fears is one of the leading risk factors for the
adoption of language legislation at the state level. Direct democracy procedures
facilitate the expression of this "grassroots fears", and states with such
provisions appear to be more likely to pass Official English laws.  likely to
pass such laws.

\section*{Statistics}

The percentage of Americans speaking languages other than English increased from
17.89\% in 2000 to 19.89\% in 2010.

\section*{Deficit}

\section*{Questions}

\section*{Political Linguistics}

\textit{Political linguistics} is a subfield of sociolinguistics that studies the ways
and extent by which governance structures influence or dictate the status and
form of language used within its jurisdiction. In this context, language becomes
a political resource over which actors can jockey for control. Once a political
choice regarding language has been made, implementation will take the form of
language planning, which can manipulate either the "corpus" or "status" of a
language \parencite{Calvet96}. Such language planning measures will be designed
by trained linguists acting in a clinical, technocratic capacity, but one must
not forget the governmental power that lies behind the measures.

\section*{Language Policy in the United States}

\cite{Baron92} and similar summaries of US language law give a solid
"legislative" history of the issue at both federal and state levels. I have not
yet reviewed \parencite{Crawford94} but I believe it will be a good access point
to primary sources. Recent activity may be less well recorded. Another problem
is with follow through. Legislation tends to have a lot of coverage at the time
of its debate directly prior to passage, when the issue is hot in the public
consciousness. Once implemented, however, it is difficult to isolate the effects
and outcomes of the policy, as the "buzz" has then moved on to new issues. This
is my fundamental goal herein: to analyse the outcomes state-level laws have
wrought for their citizens. This will be accomplished by the case studies
discussed below.

\section*{State-by-State Results}

Comparing the contents and outcomes of state langauge laws will shed light on
what form effective political mandates will take.

One interesting case is California, where voters have approved two "pro-English"
referendums in 1986 and 1998, despite the state's massive Spanish-speaking
population. It has one of the strongest Official English laws in the country.
This stance even affects the naming of babies: only the 26 letters of the
English alphabet are accepted on birth certificates, making it officially
illegal to name a child in California "José" or "María" \textcite{Larson11}.

Laws like that of Illinois sit on the other end of the spectrum, where English
is established as the official language to the extent that the
\textit{Cardinalis cardinalis} is established as the official bird, and the
Monarch butterfly as its official insect. One would expect such a language
policy to have little concrete effect.

\printbibliography

\end{document}
